---
title: "asn2139_p8105_hw3"
author: "Akanksha Nalatwad"
output: github_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width=6,
  fig.asp= .6,
  out.width= "90%"
)
theme_set(theme_minimal()+theme(legend.position="bottom"))

options(
  ggplot2.continuous.color="viridis",
  ggplot2.continuous.fill= "viridis"
)

scale_color_discrete= scale_color_viridis_d
scale_fill_discrete= scale_color_viridis_d
```


```{r}
devtools::install_github("p8105/p8105.datasets") 
library(p8105.datasets)
data("instacart")
```

How many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n>10000) %>% 
  mutate(
    aisle=factor(aisle),
    aisle= fct_reorder(aisle,n)
  ) %>% 
  ggplot(aes(x=aisle, y=n))+
  geom_point()+
  theme(axis.text=element_text(angle=90, vjust=0.5, hjust=1))
```


Let's make a table!!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetable fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank=min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle,rank) %>% 
  knitr::kable()
```
 
 Apples vs ice cream
 
```{r}

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour=mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from= order_dow,
    values_from=mean_hour
  )
```
 
## Problem 2 



#Part 1: Loading and Tidying the Dataset

```{r}
accel_df=
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440, 
               names_to="minute",
               names_prefix="activity_",
               values_to= "activity_count") %>%
                mutate(minute=as.numeric(minute)) %>%
                mutate(day2=day) %>% 
                mutate(day_type=recode(day2, "Saturday"="Weekend", "Sunday"="Weekend", "Monday"="Weekday",
                                       "Tuesday"="Weekday", "Wednesday"="Weekday", "Thursday"="Weekday",
                                       "Friday"="Weekday")) %>% 
              relocate(week,day_id,day,day_type) %>% 
          select(-day2)
               
```

Description of Dataset: This dataset follows a 65 year old man's activity count over the course of 35 days.This dataset's rows include `r nrow(accel_df)` and it's columns include `r ncol(accel_df)`.  The median activity count per minute was `r accel_df %>% pull(activity_count) %>% median()`.



## Part 2: Table of Total Activity Counts Per Day

categorized by type of day and week


```{r}

accel_modified_df=
accel_df %>% 
   select(-day_type) %>% 
   group_by(day_id,week,day) %>% 
   summarize(max_activity_per_day=max(activity_count, na.rm=TRUE), 
            mean_activity_per_day=mean(activity_count,na.rm=TRUE),
            sum_activity_per_day=sum(activity_count,na.rm=TRUE)) %>% 
   pivot_wider(names_sort=TRUE,
               names_from=day,
            values_from=sum_activity_per_day) %>% 
  relocate(day_id,week,Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, mean_activity_per_day) %>% 
  group_by(week) %>%
    summarise(Sunday= sum(Sunday, na.rm = T),
            Monday= sum(Monday, na.rm = T),
            Tuesday= sum(Tuesday, na.rm = T),
            Wednesday= sum(Wednesday, na.rm = T),
            Thursday= sum(Thursday, na.rm = T),
            Friday= sum(Friday, na.rm = T),
            Saturday=sum(Saturday,na.rm=T))
```
Apparent Trends: 
-It looks like the first week the man had less activity than any other week.
-On Saturday, the man seems to be least active since his activity counts were generally lower than on other days. 
-His most active days are Friday.


## Single-Panel Plot of 24-hour Activity Time Course
```{r}

plot_accel_df=
accel_df %>% 
      mutate(day=forcats::fct_relevel(day,c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>% 
      ggplot(aes(x=minute, y=activity_count,color=day))+
      geom_line()+ labs(title="24-Hour Activity Time Course", y="Activity count", x="Minute of Day")

ggsave("plot_accel_df.pdf", plot_accel_df, width=10, height=5)
      
```
It looks like the man starts becoming active around 6am and continues being active until he goes to sleep. His highest activiy counts occurs around 10am, 5pm, and 8pm.Generally, Sundays are generally less active days excpet around 5pm. Generally his most active days are Tuesdays.

## Problem 3


Before doing a data clean, this dataset is extremely large and bulky. It has over 5,000,000 values that follow the snowfall (mm), the depth of snow (mm), precepitation (tenths of mm), minimum temperature (in tenths of Celsius), maximum temperature(in tenths of Celsius) at different weather stations identified by id up to year 2010.There are hundreds of thousands of missing values since the weather stations are from all over the world and might not experience precipitation, or if the weather station was not able to gather data that day due to extenuating circumstances.  

##Part 1: Cleaning Datset and Finding Most Common Snowfall Values

```{r}
library(p8105.datasets)
data("ny_noaa")

ny_noaa_df=
ny_noaa %>% 
   janitor::clean_names() %>% 
  separate(date, c("year", "month", "day"), sep="-",convert=TRUE) %>% 
  mutate(prcp=as.numeric(prcp)) %>% 
  mutate(tmax= as.numeric(tmax)) %>% 
  mutate(tmin= as.numeric(tmin))

snow_df=
  ny_noaa_df %>% 
  group_by(snow) %>% 
  count() %>% 
  arrange(desc(n))
```
The most common counts of snowfall include 0 which makes sense since a lot of these weather stations might not be in areas that have either no or seasonal snowfall. The next most common count was NA, meaning those values were missing. The third most common counts of snowfall include 25 which might be an indication of some kind of cutoff that might exist when measuring/observing snowfall. 

## Two-Panel Plot

```{r}

library(ggplot2)
  ny_noaa_df %>% 
  filter(month %in% c("1","7")) %>% 
  group_by(id, year, month) %>%  
  summarize(avg_tmax=mean(tmax,na.rm=TRUE)) %>% 
  ggplot(aes(x=year, y=avg_tmax, group=id,color=id))+
  geom_point() +
geom_path() +
facet_grid(~ month)+
theme(legend.position = "none")+ labs(title="Comparing Maximum Temperatures in January vs July",
                                      y="Average Maximum Temperature (in tenths of Celsius)")
  
```

It makes sense that generally temperatures are higher during January than July. But in January it looks like the January temperatures over the years follow similar trend lines with a few outliers in 2005. In July, it looks like the July temperatures over the years follow similar trendlines with outliers in 1987, 2004, and 2010. This plot is extremely hard to intrepret since there are so many plot lines on top of each other. It would be better in the future to be able to see fewer id stations at a time.

##Two-Panel PLot

```{r}
library(patchwork)
library(tidyverse)

tmax_tmin_p=
  ny_noaa_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_hex()+ labs(title="Tmin vs Tmax", 
                   x="Minimum Temperature (in tenths of Celsius)",
                   x="Minimum Temperature (in tenths of Celsius)")
  

snowfall_p=
 ny_noaa %>% 
   janitor::clean_names() %>% 
  separate(date, c("year", "month", "day"), sep="-",convert=TRUE) %>% 
  filter(snow %in% c(1:100)) %>% 
  ggplot(aes(x=year, y=snow,group=year))+
  geom_boxplot()+ labs(title="Snowfall Histogram Over the Years", 
                       x="Year",
                       y="Snow (in mm)")

tmax_tmin_p+snowfall_p
```

